---
title: "exploring_acoustic_data"
author: "Mary Clapp"
date: "8/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(chron)
library(plotrix)


```

## POINT COUNT DATA
Median **Species Richness** per point for all surveys 2014-2016. Survey number varies by lake. First graph shows all detections within 300m of the count; the second graph shows all detections within 50m of the count.
```{r pressure, echo=FALSE}
ggplot(data = e300) +
  geom_boxplot(aes(x=fish, y=nspecies, color=fish)) +
  labs(x="treatment", y="species richness", title="Species Richness within 300m, 2014-2016") +
  facet_wrap(~basin) +
  theme_light()
```

Median **Abundance** (number of individuals counted) per point for all surveys 2014-2016.
```{r abundance, echo=FALSE}
ggplot(data = e300) +
  geom_boxplot(aes(x=fish, y=total, color=fish)) +
  labs(x="treatment", y="number of birds detected", title="Bird Abundance within 300m, 2014-2016") +
  facet_wrap(~basin) +
  theme_light()
```

Mean abundance per species per point for all surveys 2014-2016. 
```{r by-species, echo=FALSE}
ggplot(spec_stats_fish, aes(x=spec_name, y=mean_count, fill=fish)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.6) +
  geom_errorbar(position=position_dodge(),
                aes(ymin = mean_count-se_count, ymax = mean_count+se_count, width = 0.6)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Species", y = "mean count of individuals per point", title = "Mean count of individuals per point, all years")
```
I also looked at 2015 only-- no significant differences in the trends (/lack thereof).

The following graph uses only data collected on the point count at the SongMeter-- to see how congurent it is with the lake-wide data points, as well as with the microphone-collected data (next section).
```{r SM_only, eval=FALSE}
ggplot(spec_sm_stats, aes(x=species, y=mean_count, fill=fish)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.6) + 
  geom_errorbar(position=position_dodge(), 
                aes(ymin = mean_count-se_count, ymax = mean_count+se_count, width = 0.6)) +
  labs(title="Mean counts of birds within 300m of SongMeter point, 2014-2016", 
       x="Species", 
       y="mean count of individuals +/- s.e.m.")
```

### ACOUSTIC DATA
Species Richness 
```{r listening_data, eval=FALSE}
ggplot(summary, aes(x=basin, y=sp_rich, color = fish)) +
  geom_point() +
  geom_errorbar(aes(ymin = sp_rich - se, ymax = sp_rich + se), width = 0.2) + 
  labs(x = "basin", y = "average species richness", title = "Species Richness by 10-min sample")
```
Calling Activity
The following plot shows mean calling activity between fishless and fish-containing lakes. Worth noting that GCRF vocalize far more at fish-containing lakes, but were more abundant at fishless ones in the point count data.
```{r call_activity}
ggplot(foostats, aes(x=spec_name, y=mean, fill=fish)) +
  geom_bar(stat="identity", position = position_dodge(), width=0.6) +
  geom_errorbar(position=position_dodge(),
                aes(ymin = mean-sem, ymax = mean+sem, width = 0.6)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="Mean # of seconds per 10-minute recording a bird is singing", 
              x="species", 
              y="mean count of individuals +/- s.e.m.") +
  scale_fill_discrete(name="lake type")
```

